---
title: "Measurement Error Range"
author: "Victor Tsang"
date: "2023-07-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

```{r}
fossildata[[2]][!is.na(fossildata[[2]][, 2]),]
```


Task: Find out what is a typical Measurement Error variation (standard deviation) as a proportion of fossil range ($K - \theta$)

We use the fossil data provided by Cooper. First define the variables (by hand) for pulling the data from the sheet.

```{r fossildata}
library(readxl)
library(tidyverse)

sheets <- c("ArcSimAL",
  "BisPriscus.ext",
  "BisonXEur.ext+inv",
  "CervusElaEBer",
  "Coelod.ant.Brit",
  "Coelod.ant.Russ",
  "Coelod.ant.Wrang",
  "CrocCroc.Eur",
  "EquusCabEBer",
  "EquusFraEBer",
  "MammothEurCladeIII",
  "MammothPrimEBer",
  "Mammoths Eurasian",
  "MastodonNA",
  "Megaloceros",
  "NeandertalEur",
  "OvibosMos",
  "Paleolox.Japan",
  "Panth.leo.spe.Ber.Ext",
  "Panth.leo.spe.NW.ext",
  "Panth.leo.spe.OW.ext",
  "Saiga.tat.EBer",
  "Saiga.tat.Ger",
  "Ursus.spe.Ger2.ext",
  "Ursus.spe.Eur.ext")

ranges <- c("M3:N19",
  "M3:N15",
  "M4:N9",
  "M3:N59",
  "M3:N29",
  "M3:N197",
  "M3:N11",
  "M3:N81",
  "M3:N28",
  "M3:N23",
  "M3:N21",
  "M3:N36",
  "M3:N205",
  "K4:L24",
  "M3:N48",
  "M3:N147",
  "M3:N8",
  "M3:N13",
  "M3:N12",
  "M3:N25",
  "M3:N19",
  "M3:N11",
  "M3:N10",
  "M3:N17",
  "M3:N33")

sheet_ranges <- cbind(sheets, ranges)
print(sheet_ranges)
```

Next, read the excel file

```{r}

data <- lapply(1:nrow(sheet_ranges), function(i) {
  sheet_name <- sheet_ranges[i, "sheets"]
  cell_range <- sheet_ranges[i, "ranges"]
  
  # Read the specified cell range from the sheet
  sheet_data <- read_excel(path='fossildata.xlsx', sheet = sheet_name, range = cell_range, col_names = c('age', 'sd'))
  
  # Add a column for the sheet name, K, theta_hat, range, and sd_prop_range:
  sheet_data <- sheet_data %>%
    filter(!is.na(age) & !is.na(sd)) %>%
    mutate(sheet = sheet_name,
           K = max(age),
           theta_hat = min(age),
           range = K - theta_hat,
           sd_prop_range = sd / range)
  
  return(sheet_data)
})

# Combine all the data frames into a single dataframe
df <- do.call(rbind, data)

head(df)
```

Inspect the measurement error variation (sd) as a proportion of fossil record range.

```{r}
hist(df$sd_prop_range)
```

```{r}
summary(df$sd_prop_range)
```

```{r, fig.height=4}
df %>%
  ggplot(aes(x = sheet, y = sd_prop_range)) +
  geom_violin() +
  labs(x = "Sheet", y = "sd_prop_range") +
  theme_bw() +
  coord_flip()
```


## Conclusion

Generally speaking, standard deviation is a very small proportion of range. The median proportion is **3.5%**. The 75th percentile is **7.8%**. We have a very right skewed distribution (huge lengths of time and our measurement error is small relative to that.)

This is an argument for `MLE Inversion` and to a lesser extent `minmi` (regularity conditions aside). This is also an argument **against** Solow's method, which performs poorly in small measurement error scenarios.