---
title: "Scrapbook"
author: "Victor Tsang (z5209633)"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Recreating MINMI errors during simulation trials

Previously generated synthetic data is kept in `data/synthetic-data.RData`. There are two objects:

- `synthetic.data.config`: these are the constants used to generate the synthetic data.
- `datasets`: these are the synthetic datasets. For each dataset, there is a fossil record W with $n = $ `n.samples` and the error factor used. All datasets used the same set of $\sigma$'s, which are stored in `synthetic.data.config` as `fossil.sd`

```{r}
load("data/synthetic-data.RData")

synthetic.data.config
```

```{r}
head(datasets)
```


Example that works

```{r}
idx <- 3
W = as.numeric(datasets[idx, ]$W[[1]])
error_factor = as.numeric(datasets[idx, ]$error_factor[[1]])
s = synthetic.data.config$fossil.sd * error_factor


head(cbind(W, s, error_factor))
```

```{r}
rminmi::minmi(ages = W,
              sd = s,
              K = synthetic.data.config$K,
              alpha = 0.05
              # Don't give any additional tuning parameters for the Monte Carlo estimates
              # B = NA, A = NA, .B_init = NA
)
```


Example that doesnt work

```{r}
idx <- 310
W = as.numeric(datasets[idx, ]$W[[1]])
error_factor = as.numeric(datasets[idx, ]$error_factor[[1]])
s = synthetic.data.config$fossil.sd * error_factor


head(cbind(W, s, error_factor))
```

```{r}
rminmi::minmi(ages = W,
              sd = s,
              K = synthetic.data.config$K,
              alpha = 0.05
              # Don't give any additional tuning parameters for the Monte Carlo estimates
              # B = NA, A = NA, .B_init = NA
)
```
